---
title: "PPI audit slides"
author: "Dr Greig Russell"
date: "6/8/2019"
output: powerpoint_presentation
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r load_the_libraries}

library(tidyverse)
library(readxl)
library(lubridate)
library(skimr)
library(janitor)
library(scales)
library(treemap)
library(gmodels)

```

```{r load_then_tidy_the_data}
#Read the raw data and tidy the data

Raw <- read_csv("PPI.csv") %>% 
  clean_names()

Col_names <- names(Raw)
Raw_names <- as.data.frame(Col_names)
Raw_names <- mutate_at(Raw_names, vars(Col_names), str_to_title)
names(Raw) <- Raw_names$Col_names

#Remove hidden colums in the spreadsheet

Sample <- Raw %>% 
  select( Nhi, Age, Gender, Discharge_destination, Ward, Speciality, Ppi, Total_daily_dose, Duration, Started_this_admission, Concurrent_gi_irritant_therapy, Indication_recorded, If_new_on_admission_indication_in_discharge_summary, Polypharmacy)

Sample <- Sample %>% 
  rename(
    NHI = Nhi,
    PPI = Ppi)

# Load extra data fields

Raw <- read_csv("PPI_demographics.csv", na = "NULL")

Extra <- Raw %>% 
  rename(
    NHI = nhi,
    Ethnicity = ethnicitydesc1,
    Ethnicity_long = primaryethnicitybroadgroup,
    Locality = patcurrenttla
  ) %>% 
  select(NHI, Ethnicity, Ethnicity_long, Locality)

Sample <- left_join(Sample, Extra, by = "NHI")

```

```{r transform_the_data}
Factor_list <- c("Gender", "Discharge_destination", "Ward", "Speciality", "PPI", "Duration", "Started_this_admission", "Concurrent_gi_irritant_therapy", "Indication_recorded", "If_new_on_admission_indication_in_discharge_summary", "Polypharmacy", "Ethnicity", "Ethnicity_long", "Locality")

Sample <- mutate_at(Sample, vars(Factor_list), as.factor)

Sample <- Sample %>% 
  mutate(Currently_on_PPI = ifelse(!is.na(PPI), "Yes", "No"),
         Started_this_admission = fct_recode(Started_this_admission,
                                             "Yes" = "X",
                                             "Yes" = "x"),
         Indication_recorded = fct_recode(Indication_recorded,
                                          "Yes" = "yes"),
         Polypharmacy = fct_recode(Polypharmacy,
                                   "Yes" = "yes"),
         Duration = fct_recode(Duration,
                               "1 to 5"  = "1 to5"))

Sample$Currently_on_PPI <- as.factor(Sample$Currently_on_PPI)

```

## Slide 1

```{r summary}

Total_patient <- nrow(Sample)

Total_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes") %>% 
  nrow()

Percent_users = round((Total_users/Total_patient)*100, digits = 2)

```

The total number of patients in this point prevalence audit is `r Total_patient` of which `r Total_users` or `r Percent_users`% are currently prescribed a Proton Pump Inhibitor (PPI)

## Slide 2

```{r PPI_type}

Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes")

PPI_Type <- fct_count(Current_users$PPI, sort = TRUE)

treemap(PPI_Type, index = "f", vSize = "n", title = "Which is the most commonly used PPI?")

```

## Slide 3

```{r duration}

Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Duration))

Duration_count <- fct_count(Current_users$Duration, sort = TRUE)

treemap(Duration_count, index = "f", vSize = "n", title = "What has been the duration of treatment ?")

```

## Slide 4

```{r gut_irritant}

Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Concurrent_gi_irritant_therapy))

Irritant_count <- fct_count(Current_users$Concurrent_gi_irritant_therapy, sort = TRUE)

treemap(Irritant_count, index = "f", vSize = "n", title = "Where PPI users also presecibed a gut irritant medication")

```

## Slide 5

```{r dosages}
Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(PPI))

ggplot(Current_users,aes(PPI, Total_daily_dose, fill = PPI)) +
  geom_violin() +
  geom_jitter(alpha = 0.2) +
  coord_flip() +
  labs(title = "What is the total daily dose of current PPI users", x = "Total daily dose in Omeprazole equivalents", y = "PPI name") +
  guides(fill = FALSE) +
  theme_light()
```

## Slide 6

Please note "Yes" & "No" have an identical frequency, and the discontinued block is causing a visual distortion.

```{r reasons}

Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Indication_recorded))

Indication <- fct_count(Current_users$Indication_recorded, sort = TRUE)

treemap(Indication, index = "f", vSize = "n", title = "Do we know why current users are on their PPI")

```

## Slide 7

```{r discharged_summary}

Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(If_new_on_admission_indication_in_discharge_summary))

Discharge_summary <- fct_count(Current_users$If_new_on_admission_indication_in_discharge_summary, sort = TRUE)

treemap(Discharge_summary, index = "f", vSize = "n", title = "Is the PPI indication included in the discharge summary?")

```

## Slide 8

```{r polypharmacy}
PPI_poly <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Polypharmacy))

PPI_poly_count <- fct_count(PPI_poly$Polypharmacy, sort = TRUE)

treemap(PPI_poly_count, index = "f", vSize = "n", title = "In PPI users, is there also polypharmacy present?")
```

## Slide 9

```{r age}

ggplot(Sample, aes(Age, fill = Currently_on_PPI)) +
  geom_bar(binwidth = 5) +
  labs(title = "What is the age profile for PPI & non-PPI users?")

```

## Slide 10
 
```{r age_density}
ggplot(Sample, aes(Age, color = Currently_on_PPI)) +
  geom_density() +
  labs(title = "Is this age difference statistically significant", subtitle = "Answer = Yes, p = 0.002")
```
 
## Slide 11
 
```{r gender}

Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Gender))

Gender_count <- fct_count(Current_users$Gender, sort = TRUE)

treemap(Gender_count, index = "f", vSize = "n", title = "Gender balance in PPI user")

```
 
## Slide 12

No significant difference is seen (p = 0.16)
```{r speciality}

Table_chart <- Sample %>% 
  rename(Specialty = Speciality) %>% 
  tabyl(Specialty, Currently_on_PPI) %>% 
  adorn_totals(where = "row") %>% 
  adorn_percentages(denominator = "col") %>% 
  adorn_pct_formatting(digits = 0) %>% 
  adorn_ns(position = "front")

kable(Table_chart)

```

## Slide 13

No significant difference is seen (p = 0.24)
```{r ward}

Table_chart <- Sample %>% 
  tabyl(Ward, Currently_on_PPI) %>% 
  adorn_totals(where = "row") %>% 
  adorn_percentages(denominator = "col") %>% 
  adorn_pct_formatting(digits = 0) %>% 
  adorn_ns(position = "front")

kable(Table_chart)
```

## Slide 14

No significant difference is seen (p = 0.1)
```{r locality}

Table_chart <- Sample %>% 
  tabyl(Locality, Currently_on_PPI) %>% 
  adorn_totals(where = "row") %>% 
  adorn_percentages(denominator = "col") %>% 
  adorn_pct_formatting(digits = 0) %>% 
  adorn_ns(position = "front")

kable(Table_chart)

```

## Slide 15

No Significant difference is seen (p = 0.21)
```{r ethnicity}
Table_chart <- Sample %>% 
  tabyl(Ethnicity, Currently_on_PPI) %>% 
  adorn_totals(where = "row") %>% 
  adorn_percentages(denominator = "col") %>% 
  adorn_pct_formatting(digits = 0) %>% 
  adorn_ns(position = "front")

kable(Table_chart)

```

## Slide 16

No significant difference is seen (p=0.29)
```{r discharge_destination}

Table_chart <- Sample %>% 
  tabyl(Discharge_destination, Currently_on_PPI) %>% 
  adorn_totals(where = "row") %>% 
  adorn_percentages(denominator = "col") %>% 
  adorn_pct_formatting(digits = 0) %>% 
  adorn_ns(position = "front")

kable(Table_chart)
```

