---
title: "PPI analysis"
output: html_notebook
---

```{r}
# Load the tidyverse

library(tidyverse)

```

```{r}
# Load other libraries

library(readxl)
library(lubridate)
library(skimr)
library(janitor)
library(scales)
library(treemap)
library(gmodels)

```

```{r}
#Read the raw data and clean up the names

Raw <- read_csv("PPI.csv") %>% 
  clean_names()

Col_names <- names(Raw)
Raw_names <- as.data.frame(Col_names)
Raw_names <- mutate_at(Raw_names, vars(Col_names), str_to_title)
names(Raw) <- Raw_names$Col_names

#Remove hidden colums in the spreadsheet

Sample <- Raw %>% 
  select( Nhi, Age, Gender, Discharge_destination, Ward, Speciality, Ppi, Total_daily_dose, Duration, Started_this_admission, Concurrent_gi_irritant_therapy, Indication_recorded, If_new_on_admission_indication_in_discharge_summary, Polypharmacy)

Sample <- Sample %>% 
  rename(
    NHI = Nhi,
    PPI = Ppi)

glimpse(Sample)
```

```{r}
# Load extra data fields

Raw <- read_csv("PPI_demographics.csv", na = "NULL")

Extra <- Raw %>% 
  rename(
    NHI = nhi,
    Ethnicity = ethnicitydesc1,
    Ethnicity_long = primaryethnicitybroadgroup,
    Locality = patcurrenttla
  ) %>% 
  select(NHI, Ethnicity, Ethnicity_long, Locality)

glimpse(Extra)
```

```{r}
# Join the extra fields to the Sample data

Sample <- left_join(Sample, Extra, by = "NHI")

glimpse(Sample)
```

```{r}
# Transform the data for analysis

Factor_list <- c("Gender", "Discharge_destination", "Ward", "Speciality", "PPI", "Duration", "Started_this_admission", "Concurrent_gi_irritant_therapy", "Indication_recorded", "If_new_on_admission_indication_in_discharge_summary", "Polypharmacy", "Ethnicity", "Ethnicity_long", "Locality")

Sample <- mutate_at(Sample, vars(Factor_list), as.factor)

glimpse(Sample)
```



```{r}
#Code chunckj to extract NHI data to request additional data fields from DA - thanks to Paul & Quentin for providing additional data.

# DA_Data <- Raw %>% 
#   select(NHI)
# write.table(DA_Data, "/home/greig/R-projects/PPI_2019/PPI_NHI.csv", row.names = FALSE)
```

```{r}
# Load the population data

Raw <- read_csv("PPI_Population.csv")
```
```{r}
Population <- Raw %>% 
  filter(year(DDATE) == 2018) %>% 
  select(ADMAGE, GENDER, ETHGROUP, DOMTLA, SPECIALTY)

Col_names <- names(Population)
Raw_names <- as.data.frame(Col_names)
Raw_names <- mutate_at(Raw_names, vars(Col_names), str_to_title)
names(Population) <- Raw_names$Col_names

Population <- Population %>% 
  rename(Age = Admage,
         Ethnicity = Ethgroup,
         Locality = Domtla,
         Speciality = Specialty
         )

Population <- Population %>% 
  mutate(Gender = fct_recode(Gender,
                             "Male" = "M",
                             "Female" = "F"))

Factor_list <- c("Ethnicity", "Locality", "Speciality")
Population <- mutate_at(Population, vars(Factor_list), as.factor)

glimpse(Population)

```

```{r}
# Further transformations of the Sample data set for ease of anaylsis

Sample <- Sample %>% 
  mutate(Currently_on_PPI = ifelse(!is.na(PPI), "Yes", "No"),
         Started_this_admission = fct_recode(Started_this_admission,
                                             "Yes" = "X",
                                             "Yes" = "x"),
         Indication_recorded = fct_recode(Indication_recorded,
                                          "Yes" = "yes"),
         Polypharmacy = fct_recode(Polypharmacy,
                                   "Yes" = "yes"),
         Duration = fct_recode(Duration,
                               "1 to 5"  = "1 to5"))

Sample$Currently_on_PPI <- as.factor(Sample$Currently_on_PPI)

```

# PPI usage within PNH - a shapshot
This study is the point prevalence of PPI usage within PNH. The basic parameters of the study are;

```{r}
Total_patient <- nrow(Sample)
Total_patient

Total_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes") %>% 
  nrow()
Total_users

Percent_users = round((Total_users/Total_patient)*100, digits = 2)
#scales::percent(Percent_users)
Percent_users
```

Of those folks on a PPI which PPI is most common;

```{r}
Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes")

PPI_Type <- fct_count(Current_users$PPI, sort = TRUE)

PPI_Type

treemap(PPI_Type, index = "f", vSize = "n", title = "Which PPI is used most commonly")
```

How long had they been on the PPI
```{r}
Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Duration))

Duration_count <- fct_count(Current_users$Duration, sort = TRUE)

Duration_count

treemap(Duration_count, index = "f", vSize = "n", title = "How long have users been taking their PPI")
```

Of patients currently on PPI, how many are also on GI irritant medication
```{r}
Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Concurrent_gi_irritant_therapy))

Irritant_count <- fct_count(Current_users$Concurrent_gi_irritant_therapy, sort = TRUE)

Irritant_count

treemap(Irritant_count, index = "f", vSize = "n", title = "Where patients on PPI also on GI irritant medication")
```

Of patients currently on PPI what is their total daily dose in Omeprazole equivalents
```{r}
Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(PPI))

ggplot(Current_users,aes(PPI, Total_daily_dose, fill = PPI)) +
  geom_violin() +
  geom_jitter(alpha = 0.2) +
  coord_flip() +
  labs(title = "What is the total daily dose of current PPI users") +
  guides(fill = FALSE) +
  theme_light()
```

Do we know why current users of PPI are on this medication
```{r}
Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Indication_recorded))

Indication <- fct_count(Current_users$Indication_recorded, sort = TRUE)

Indication

treemap(Indication, index = "f", vSize = "n", title = "Do we know why current users are on their PPI")
```

If the PPI was started on this admission was the indication in the discharge summary
```{r}
Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(If_new_on_admission_indication_in_discharge_summary))

Discharge_summary <- fct_count(Current_users$If_new_on_admission_indication_in_discharge_summary, sort = TRUE)

Discharge_summary

treemap(Discharge_summary, index = "f", vSize = "n", title = "Does the discharge summary tell why the patient was started on PPI")
```

Is polypharmacy common in patients on PPI
```{r}
PPI_poly <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Polypharmacy))

PPI_poly_count <- fct_count(PPI_poly$Polypharmacy, sort = TRUE)

PPI_poly_count

treemap(PPI_poly_count, index = "f", vSize = "n", title = "Proportion of PPI users also with Polypharmacy")
```

## Who are the PPI users as compared other patients currently in hospital

Is the age profile of patients on PPI different from others
```{r}

ggplot(Sample, aes(Age, fill = Currently_on_PPI)) +
  geom_bar(binwidth =5) +
  labs(title = "Is the Age profile different between PPI & non-PPI users?")

ggplot(Sample, aes(Age, color = Currently_on_PPI)) +
  geom_density() +
  labs(title = "Is there a difference between the age profile of PPI users and non-users", subtitle = "Answer = Yes, p = 0.002")

Age_1 <- Sample %>% 
  filter(Currently_on_PPI == "Yes")

Age_2 <- Sample %>% 
  filter(Currently_on_PPI == "No")

t.test(Age_1$Age, Age_2$Age)
```

Is there a gender difference between users and non-users
```{r}
CrossTable(Sample$Currently_on_PPI, Sample$Gender, prop.r = FALSE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

Current_users <- Sample %>% 
  filter(Currently_on_PPI == "Yes",
         !is.na(Gender))

Gender_count <- fct_count(Current_users$Gender, sort = TRUE)

Gender_count

treemap(Gender_count, index = "f", vSize = "n", title = "Gender balance for PPI users")
```

### Is there any difference by Speciality or Ward?
The answer is no (Speciality p = 0.16, Ward p = 0.24)
```{r}
CrossTable(Sample$Currently_on_PPI, Sample$Speciality, prop.r = FALSE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

CrossTable(Sample$Currently_on_PPI, Sample$Ward, prop.r = FALSE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
```

### Is there any difference between locality or ethnicity? 
The answer is no (Locality p = 0.1, Ethnicity p = 0.21)
```{r}
CrossTable(Sample$Currently_on_PPI, Sample$Locality, prop.r = FALSE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

CrossTable(Sample$Currently_on_PPI, Sample$Ethnicity, prop.r = FALSE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
```

### Does the discharge destination have any impact on PPI usage?
The answer is no (p = 0.29).
```{r}
CrossTable(Sample$Currently_on_PPI, Sample$Discharge_destination, prop.r = FALSE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

```

